The genomic impacts of drift and selection for hybrid performance in maize

### 2015-08-19 main.tex
revisions to intro from Jode/Justin

Multiple studies with molecular markers have indicated that these heterotic groups have diverged genetically over time to become highly structured and isolated populations, resulting in a dramatic restructuring of population genetic variation \citep{duvick2004long, ho2005extent, feng2006temporal}. 
Advances in high throughput genotyping and the development of a maize reference genome now enable the observation of maize population structure at high marker density across the whole genome \citep{ganal2011a-large,chia2012maize}. 
These studies have examined a broad spectrum of germplasm at various points in the history of maize to search for the signals of population structure and artificial selection \citep{Hufford2012b, van2012historical,jiao2012genome}. 
Although selective sweeps remaining from domestication are clearly visible, the impact of selection during modern breeding appears comparatively small in terms of its impact on genomic diversity \citep{Hufford2012b,van2012historical}, despite steady, heritable improvement in phenotype \citep{duvick2005contribution}. 
The lack of distinct selection signals from modern breeding may be due to population-specific selection; this interpretation is supported by the measured success in identifying targets of selection in individual experimental populations selected for a limited number of phenotypes \citep{hirsch2014insights, beissinger2014genome}.

Here we take a detailed look at the dynamics of genome-wide patterns of diversity over time within an individual breeding program. 
The Iowa Stiff Stalk Synthetic (BSSS) and the Iowa Corn Borer Synthetic No. 1 (BSCB1) Reciprocal Recurrent Selection Program of the USDA-ARS at Ames, Iowa (hereafter referred to as the Iowa RRS) represents one of the best-documented public experiments on selection for combining ability and hybrid performance. 
BSSS and BSCB1 have been recurrently selected for improved cross-population hybrids \citep{penny1971twenty}. 
\jri{Jode/Justin add text explaining recurrent selection and populations}
This model of selection, named reciprocal recurrent selection, provides the generalized model for strategies commonly used in commercial maize breeding \citep{comstock1949breeding, duvick2004long}. 
The Iowa RRS experiment proves especially relevant because lines derived from the BSSS population have had a major impact upon the development of commercial hybrids \citep{duvick2004long, darrah1986}, the formation of modern heterotic groups \citep{troyer1999background, senior1998utility}, and the choice of a maize reference genome \citep{schnable2009the-b73-maize}.
\jri{something about heterosis}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% INTRO

>>>>

\rev{Although the founders of these heterotic groups were not initially differentiated, m}ultiple studies with molecular markers have indicated that these heterotic groups have diverged genetically over time to become highly structured and isolated populations, resulting in a dramatic restructuring of population genetic variation \citep{duvick2004long, ho2005extent, feng2006temporal}. 
Advances in high throughput genotyping and the development of a maize reference genome now enable the observation of maize population structure at high marker density across the whole genome \citep{ganal2011a-large,chia2012maize}. 
These studies have examined a broad spectrum of germplasm at various points in the history of maize to search for the signals of population structure and artificial selection \citep{Hufford2012b, van2012historical,jiao2012genome}. 
Although selective sweeps remaining from domestication are clearly visible, the impact of selection during modern breeding appears comparatively small in terms of its impact on genomic diversity \citep{Hufford2012b,van2012historical}, despite steady, heritable improvement in phenotype \citep{duvick2005contribution}. 
The lack of distinct selection signals \rev{in broad germplasm collections may be due to a population specific selection within germplasm subgroups, which would be further obscured by migration due to the exchange of material among breeders.} \textcolor{red}{cite Mikel here}  \rev{In this case selection will be easier to detect in populations maintained in an individual program.} This \rev{possibliity} is supported by the measured success in identifying targets of selection in individual experimental populations \rev{under directional selection for specific phenotypes such as seed size and ear number.} \citep{hirsch2014insights, beissinger2014genome}.

\rev{In this study, we examine a different closed experiment whose phenotypic target and selection method -Reciprocal Recurrent Selection- closely mirrors methods used in the generation of modern maize hybrids \citep{comstock1949breeding, duvick2004long}. Proposed by Comstock et al. (1949), reciprocal recurrent selection is a procedure to simultaneously improve two populations for their ability to produce high yielding hybrids when individuals from the two populations are crossed. In this method, the two populations are kept separate and lines from each population are advanced to the next cycle of selection based on the performance of the line or its close relatives in a testcross with random individuals from the opposing population. Selection is therefore carried out on lines within each population, but on a cross-population hybrid phenotype. After selection the selected lines within each population are intermated to generate lines for the next cycle. In this way the two populations simultaneously evolve improved combining ability with one another. The USDA-ARS at Ames, Iowa has conducted a Reciprocal Recurrent Selection Program (hereafter referred to as the Iowa RRS) with the Iowa Stiff Stalk Synthetic (BSSS) and Iowa Corn Borer Synthetic No. 1 (BSCB1) populations for 18 cycles of selection\citep{penny1971twenty}} \textcolor{red}{cite Jode 2011 here}. \rev{This program represents one of the best-documented public experiments on selection for combining ability and hybrid performance. Reminiscent of modern maize germplasm as a whole, BSSS and BSCB1 have evolved from undifferentiated founders as they have been selected for cross-population hybrid yield. BSSS and BSCB1 have both been maintained as closed populations with no intentional migration (although evidence suggests low level of unintentional gene flow early in the BSSS population). Extensive records on population sizes, breeding methods, selection differentials, and seed are available for all cycles of selection. This set of resources on a closed population selected for hybrid yield make the Iowa RRS an ideal test case for the study of the genomic impact of intensive hybrid breeding in maize.} The Iowa RRS experiment \rev{provides additional historical relevance} because lines derived from the BSSS population have had a major impact upon the development of commercial hybrids \citep{duvick2004long, darrah1986}, the formation of modern heterotic groups \citep{troyer1999background, senior1998utility}, and the choice of a maize reference genome \citep{schnable2009the-b73-maize}.
\jri{something about heterosis} \jpg{what are you thinking for heterosis here?}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% INTRO
